Forecasting Gamma-ray Bursts using Gravitational Waves

### 2018-02-24 Paper/Manuscript.tex
Started writing Sec VI

\documentclass[amsmath,amssymb,aps,floats,amsfonts,notitlepage,superscriptaddress,eqsecnum,nofootinbib,10pt]{revtex4-1}
\linespread{1.2}
%\usepackage{booktabs} % for tabular improvments
\usepackage{graphicx}       % Include figure files
\usepackage{dcolumn}        % Align table columns on decimal point
\usepackage{bm}             % bold math
\usepackage{tensor}
\usepackage{url}
\usepackage[colorlinks]{hyperref}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{float}
\usepackage[usestackEOL]{stackengine} % package used to write several lines in one box
\allowdisplaybreaks[1]

@@

It remains to compute $|Q(\theta,\phi,\psi,\iota)|^2$. 
Since we can not know the direction of the source a priori, it is appropriate to average over the angles $\theta,\phi,\psi,\iota$ which we compute
%For our computations, we will only work with one detector so we can set $\psi=0$ without loss of generality, which results in
%%
%
%\be
%F_+(\theta,\phi) = \f{1}{2}(1+\cos^2\theta)\cos2\phi, \qquad F_\times(\theta,\phi) = \cos\theta \sin2\phi.
%\ee
%
%%
via the following RMS-averaging integral
%%
%
\be
\langle F^2_{+}\rangle \equiv \f{1}{2\pi}\int_0^{2\pi}d\psi \,\f{1}{4\pi}\int_0^{2\pi} d\phi\,\int_0^\pi d\theta \sin\theta\ F^2_{+}(\theta,\phi,\psi)=\f{1}{5}\label{eq:ang_av}
\ee
%
%%
and likewise $\langle F^2_{\times}\rangle=1/5 $.
Finally, we average over the orbital inclination angle
%
\be
%\langle g(\iota) \rangle = 
\f{1}{2} \int_0^\pi d\iota\sin\iota\left[\left(\f{1+\cos^2\iota}{2}\right)^2+\cos^2\iota\right] = \f{4}{5} .\label{eq:inc_ang_av}
\ee
Thus we have that
%
\be
\langle |Q(\theta,\phi,\psi,\iota)|^2 \rangle_{\theta,\phi,\psi,\iota} = \f{4}{25}\label{eq:RMS_th_ph_ps_iota}.
\ee
%
As we will also consider the case of $\psi=0,\iota=0$, we additionally introduce the following $\theta,\phi$-only averages
%
\be
 \langle F_+^2\rangle_{\theta,\phi} = \frac{7}{30},\qquad \langle F_\times^2\rangle_{\theta,\phi} = \frac{1}{6}
\ee
yielding
%
\be
\langle |Q(\theta,\phi,\psi=0,\iota=0)|^2 \rangle_{\theta,\phi} = \f{2}{5} \label{eq:RMS_th_phi}.
\ee
Finally, we add the case of an optimally-oriented, purely $+$-polarized source ($\psi=0,\iota=0$) that is directly overhead ($\theta=0$) simply resulting in 
%
\be
\langle |Q(\theta=0,\phi=0,\psi=0,\iota=0)|^2 \rangle = 1.
\ee
%
%
%
%\be
%\tilde{H}(f)\equiv  A\,\left\langle\, |Q(\theta,\phi,\psi,\iota)|^2\,\right\rangle^{1/2}\, \f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6} f^{-7/6}\, . \label{eq:ang_av_strain_FD}
%\ee
%

In a network of $N$ interferometers, we can define a total effective SNR as follows
%%
%
\be
\rho_\text{tot} = \left[{\sum_{i=1}^N \rho^2_i}\right]^{1/2} \label{eq:SNR_total}
\ee
%
%%
where Eq.~(\ref{eq:SNR_v2}) is employed to compute $\rho_i$ coming from individual interferometers each with a specific quality factor $Q_i$.
Here, instead of working with specific quality factors of the LIGO and the Virgo interferometers, we consider
an \emph{imaginary network} of three detectors ($N=3$), where one interferometer is optimally oriented/positioned, another one operates at the RMS average thus
its sensitivity is relatively suppressed by a factor $2/5$ (Eq.~\ref{eq:RMS_th_ph_ps_iota}) due to suboptimal configuration.
Our third detector has even lower sensitivity with a relative suppression factor of $4/5-\sqrt{2/5}\simeq 0.1675$ which is essentially a $\{\theta$-$\phi\}$-RMS away
from the overall RMS averaged value of 2/5.
This is our way of simulating a future detection in the Advanced-LIGO-Virgo network of 2020s where the three interferometers will yield varying SNRs
due to their differing quality factors.
%
Therefore, to represent our multi-interferometer network, we introduce the norm of the frequency-domain strain (\ref{eq:strain_FD}) for each interferometer
%%
%
\be
\tilde{H}_i(f)\equiv \tilde{A}_i\, h_0\, f^{-7/6}, \label{eq:Hi_s}
\ee
%
%%
where
%%
%
\begin{align}
%\tilde{A}_i &\equiv A \langle |Q|^2\rangle^{1/2}_i=\f{1}{2}\sqrt{\f{5}{6}}\, \pi^{-2/3} \left\langle\, |Q_i(\theta,\phi,\psi,\iota)|^2\,\right\rangle^{1/2}\quad\text{and} \label{eq:A_tilde}\\
h_0 &= \f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6} \label{eq:h0}
\end{align}
%
and the $\tilde{A}_i$ are given in Table~\ref{table:network_params}. %for $i=1,2,3$. %Clearly, the strain and thus the SNR at each of our imaginary interferometers scales as $\langle |Q|^2\rangle^{1/2}_i$, thus
%%
The SNR for each interferometer can now be written as
%%
%
\begin{align}
\rho_i &= 2\tilde{A}_i\, h_0 \left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{n}(f)} \right]^{1/2}\label{eq:SNR_i}.
\end{align}
%
We arbitrarily make the following assignments for our network: $h=1, m=2, l=3$ representing high, medium, low strains/SNRs, respectively. 
We summarize our network in the following Table~\ref{table:network_params}.
%%

\begin{table}[h]
\centering
\begin{tabular}{c||c|c|l}
 IFO ($i$) & configuration %& $\langle |Q_i|^2\rangle$ 
 & $\tilde{A}_i$ & $\tilde{H}_i(f)$\\
 \hline
 \hline
 1 & high ($h$) &  A & $A h_0\, f^{-7/6}$ \T\\
 2 & medium ($m$) & $\tfrac{2}{5}A$ & $ \tfrac{2}{5}A h_0 \, f^{-7/6}$\B\T \\
 3 & low ($l$)  & $\left(\tfrac{4}{5}-\sqrt{\tfrac{2}{5}}\right)A$ & $ \left(\tfrac{4}{5}-\sqrt{\tfrac{2}{5}}\right)A h_0 \, f^{-7/6}$\B \\
 \hline
\end{tabular}
\caption{%The averaged quality factors $\langle|Q|^2\rangle$ of 
Our imaginary network of three L-shaped interferometers (IFOs). Column four lists the magnitude of the strain for each IFO via Eq.~(\ref{eq:Hi_s}).
``Configuration'' describes how ideally oriented/positioned an IFO is with respect to a compact binary source. High means ideal configuration maximizing $Q_i$.
Medium is the RMS-averaged response. Recall $A= \pi^{-2/3} \sqrt{5/24}$. 
%Note that for an interferometer where the angle between the arms is $\alpha$, $A$ rescales to $A\sin\alpha$.
}\label{table:network_params}
\end{table}
%
%
We can now write the total SNR for our network
%%
%
\begin{align}
\rho_\text{tot} &= \sqrt{\rho^2_h+\rho^2_m+\rho^2_l} \label{eq:SNR_total_3IFOs} .
%	      &= 2A h_0 \sqrt{1+\tfrac{2}{5}+\tfrac{4}{25}} \left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{noise}(f)}\right]^{1/2}\nn\\
%	      &= \sqrt{\frac{13}{10}}\,\pi^{-2/3}\,\f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6}\left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{noise}(f)}\right]^{1/2} \label{eq:SNR_total_3IFOs_v2}.
\end{align}
%
%%
%Although we made several simplifying assumptions to obtain $\rho_\text{tot}$, these should not affect our total SNR estimation too much.
%For example, averaging over all the angles underestimates the SNR of an interferometer with a near-ideal orientation with respect to the incoming GWs
%while also overestimating the SNR of a badly oriented detector. Rewriting Eq.~(\ref{eq:SNR_total}) as Eq.~(\ref{eq:SNR_total_3IFOs}) compensates
%for these.
We will start Sec.~\ref{sec:results} by testing our network with GW170817 whose signal got picked up by the LIGO detectors around $f=30$\,Hz.
So we will change the lower limit of the integral in Eq.~(\ref{eq:SNR_total_3IFOs}) from $0$ to $30$.
For our advance warning times we will further use separate initial frequencies for each inteferometer as we will explain in Sec.~\ref{sec:results}.

@@

The amplitude-adjusted strains $ \tilde{H}_i(f)$ are represented by the solid ($i=h$), dashed ($i=m$), and dotted ($i=l$) lines in the figure where we actually plot
$2\sqrt{f} \tilde{H}_i(f)$ as explained above.

@@

\bar\rho_h = 35.8,\quad \bar\rho_m = 14.3,\quad \bar\rho_l = 6.00 \label{eq:GW170817_SNRs}.

@@

To accommodate more realistic orientations, we show two shaded regions in Fig.~\ref{fig:figGW170817}. The inspirals in the lighter region yield SNR$\ge 22.34$ as
they sweep from 10\,Hz to $f_\text{ISCO} \simeq1605\,$Hz whereas inspirals in the darker region accumulate $ \bar\rho_l \le \text{SNR} \le 22.34$.
Given a network of three equally sensitive, randomly oriented IFOs we expect that a single BNS inspiral would produce two detector strains in the darker region 
and one in the lighter region.

@@

\caption{GW170817 as it sweeps from 10\,Hz to $f_\text{ISCO} \simeq1605\,$Hz across our network with ASD (sensitivity) of LIGO Livingstone (L1) during the event (red curve).
We also show the Hanford detector's sensitivity for comparison. The dashed, solid and dotted lines represent the strains 
$2\sqrt{f} \tilde{H}_i(f)$ (see, e.g,. Eq.~(\ref{eq:Hi_s}) and Table~\ref{table:network_params}) due to the inspiral of
a binary neutron star system with parameters given in Eq.~(\ref{eq:GW170817_params}). The change in strain amplitudes is the result of a
given interferometer's orientation and the sky position of the source. In other words, the dashed line is how ``loud'' GW170817 
is at an optimally oriented IFO with L1's sensitivity

@@

(quality factor $Q=1$). Similarly, the solid line is the strain in a randomly oriented IFO with randomized source sky location, hence
we use the RMS-averaged L1 sensitivity which effectively lowers the solid line by a factor of $2/5$. %, but with RMS-average over all angles 

@@

resulting in a reduction of amplitude to $ 16.7\%$ of the optimal case. 

@@

network, we expect all three strains to exist somewhere in the shaded region.

@@

We can now compute the total SNR of GW170817 as inferred by our network
%%
%
\be
\bar\rho_\text{tot}= \sqrt{\bar\rho^2_h+\bar\rho^2_m+\bar\rho^2_l}= 39.1 \, \label{eq:SNR_total_GW170817} .
\ee
%
%%
This also compares well with the actual value of 32.4. Therefore, we conclude that our imaginary network performs realistically enough to make predictions on the performance of IFOs in the 2020s and 2030s.
%
%
%
%
%
\section{Results: Advance warning times for inspiralling neutron stars}\label{sec:results}
 %\subsection{BNS inspirals at Advanced LIGO-Virgo and the Einstein Telescope}\label{Sec:BNS_2020_2030}
 %Recall that we are interested in the advance warning times that the (near)-future ground-based interferometers can provide for the optical astronomers before the prompt GRB at the end of the merger. 
 Our aim is to foretell GRBs using our network made up of three IFOs with A-LIGO design sensitivity in the 2020s and the Einstein Telescope sensitivity in the 2030s.
 Our predictive capabilities depend on how much advance warning the future networks will give us prior to the prompt merger/GRB.
 Our definition of advance warning is the inspiral time to the merger from a certain threshold instant $\bar{t}$ at which point the network ``agrees'' that there is an inspiral 
 event statistically significant enough to issue warnings to electromagnetic telescopes.
 We define this threshold instant to be given at a frequency $\bar{f}$ when the total SNR equals 15. 
 More specifically, in the case of the Advanced-LIGO-Virgo network (ALV), our statement is that there exists a frequency $\bar{f} < f_\text{ISCO}$ such that
 %
 \be
 \bar\rho_\text{tot}= \sqrt{\left[\rho_h(f_0^h,\bar{f})\right]^2+\left[\rho_m(f_0^m,\bar{f})\right]^2+\left[\rho_l(f_0^l,\bar{f})\right]^2}=15, \label{eq:SNR_bar}%2\tilde{A}_m\, h_0\left[ \int_{f_0}^{{\bar{f}}} df'\, \f{f'^{-7/3}}{S_\text{noise}(f')}\right]^{1/2} =15 .
 \ee
 %
 where $f_0^i$ are given by the solution set to
 %there exists some frequency $f_0 < \bar{f}$ such that
 \be
 \sqrt{S_\text{n}(f_0^i)} = 2\sqrt{f_0^i}\, \tilde{H}_i(f_0^i) \label{eq:f0},\qquad i=h,m,l
 \ee
 %
 with $f_0^i < \bar{f}$. In other words, $f_0^i$ are the initial GW frequencies at which a given inspiral enters the $i$th IFO's sensitivity band. 
 Once we know $\bar{f}$, we can compute the remaining time to the merger, $\tau_\text{insp}(\bar{f})$, using Eq.~(\ref{eq:tau_insp})
 and its 3.5PN-enhanced version from Ref.~\cite{Blanchet_LRR}. This is what we call our advance warning time $T_\text{AW}$.
 For simplicity, we set $m_1=m_2=1.4 M_\odot \implies M_c\simeq 1.219 M_\odot$, which we justified in
 Point 3 of Sec.~\ref{sec:idealizations}. 
 With the masses fixed, the only variable determining $h_0$ is the luminosity distance $D$ as can be seen from Eq.~(\ref{eq:h0}).
 We summarize our procedure in the flow diagram below.
 %
 %
  \begin{figure}[h!]
 \framebox{\Longstack[l]{Pick $D$}}$\longrightarrow$\framebox{\Longstack[l]{Construct\\ $\tilde{H}_i(f)$}} 
 $\longrightarrow$  \framebox{\Longstack[l]{Compute $f_0^i$\\ from Eq.(\ref{eq:f0})}} $\longrightarrow$  \framebox{\Longstack[l]{Compute $\bar{f}$\\ from Eq.(\ref{eq:SNR_bar})}}
 $\longrightarrow$  \framebox{\Longstack[l]{Compute $T_{AW}=\tau_\text{insp}(\bar{f})$\\ from Eq.~(\ref{eq:tau_insp})}}
  \caption{Flow diagram summarizing our procedure for computing advance warning times $T_\text{AW}$ with the Advanced-LIGO-Virgo network. 
  We will compute $T_\text{AW}$ for the case of $M=2.8M_\odot$ BNS system inspiralling at $D=50,100,200\,$Mpc. 
  %for A-LIGO and $D=200,400,1000\,$Mpc for the ET.
  }\label{fig:flow}
\end{figure}

@@

\section{Binary black hole neutron star systems}
In this section, we investigate the possibility of witnessing the tidal distruption of a neutron star as it falls into a black hole. 
The treatment of the inspiral is the same as before, however the fate of the neutron star depends strongly on the system's intrinsic parameters
such as the black hole mass and spin. The evolution of black hole-neutron star (BH-NS) binaries is a very active branch of numerical relativity
(see Ref. \cite{Shibata:2011jka} for a comprehensive review).

>>>>

\documentclass[amsmath,amssymb,aps,floats,amsfonts,notitlepage,superscriptaddress,eqsecnum,nofootinbib,10pt]{revtex4-1}
\linespread{1.2}
%\usepackage{booktabs} % for tabular improvments
\usepackage{graphicx}       % Include figure files
\usepackage{dcolumn}        % Align table columns on decimal point
\usepackage{bm}             % bold math
\usepackage{tensor}
\usepackage{url}
\usepackage[colorlinks]{hyperref}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{float}
\usepackage{enumerate}
\usepackage[usestackEOL]{stackengine} % package used to write several lines in one box
\allowdisplaybreaks[1]

@@

It remains to compute $|Q(\theta,\phi,\psi,\iota)|^2$. 
Since we can not know the direction of the source a priori, it is appropriate to average over the angles $\theta,\phi,\psi,\iota$ which we compute
%For our computations, we will only work with one detector so we can set $\psi=0$ without loss of generality, which results in
%%
%
%\be
%F_+(\theta,\phi) = \f{1}{2}(1+\cos^2\theta)\cos2\phi, \qquad F_\times(\theta,\phi) = \cos\theta \sin2\phi.
%\ee
%
%%
via the following RMS-averaging integral
%%
%
\be
\langle F^2_{+}\rangle \equiv \f{1}{2\pi}\int_0^{2\pi}d\psi \,\f{1}{4\pi}\int_0^{2\pi} d\phi\,\int_0^\pi d\theta \sin\theta\ F^2_{+}(\theta,\phi,\psi)=\f{1}{5}\label{eq:ang_av}
\ee
%
%%
and likewise $\langle F^2_{\times}\rangle=1/5 $.
Finally, we average over the orbital inclination angle
%
\be
%\langle g(\iota) \rangle = 
\f{1}{2} \int_0^\pi d\iota\sin\iota\left[\left(\f{1+\cos^2\iota}{2}\right)^2+\cos^2\iota\right] = \f{4}{5} .\label{eq:inc_ang_av}
\ee
Thus we have that
%
\be
\langle |Q(\theta,\phi,\psi,\iota)|^2 \rangle_{\theta,\phi,\psi,\iota} = \f{4}{25}\label{eq:RMS_th_ph_ps_iota}.
\ee
%
As angles have zero averages in general, we can think of the RMS as the standard deviation ($\sigma =2/5$). 
Therefore, $2\sigma= 4/5$ above the average (zero in this case) should statistically cover 95\% of the four-dimensional angle space. 
%Finally, we add the case of an optimally-oriented, purely $+$-polarized source ($\psi=0,\iota=0$) that is directly overhead ($\theta=0$) simply resulting in 
%
%\be
%\langle |Q(\theta=0,\phi=0,\psi=0,\iota=0)|^2 \rangle = 1.
%\ee
%
%
%
%\be
%\tilde{H}(f)\equiv  A\,\left\langle\, |Q(\theta,\phi,\psi,\iota)|^2\,\right\rangle^{1/2}\, \f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6} f^{-7/6}\, . \label{eq:ang_av_strain_FD}
%\ee
%
%%

In a network of $N$ interferometers, the total SNR is defined as follows
%%
%
\be
\rho_\text{tot} = \left[{\sum_{i=1}^N \rho^2_i}\right]^{1/2} \label{eq:SNR_total}
\ee
%
%%
where Eq.~(\ref{eq:SNR_v2}) is employed to compute $\rho_i$ coming from individual interferometers each with a specific quality factor $Q_i$.
Here, instead of working with specific quality factors of the LIGO and the Virgo interferometers, we consider
an \emph{model network} of three detectors ($N=3$), where one interferometer is near-optimally oriented/positioned thus it is at $2\sigma$ above the average.
The second IFO operates at the RMS average; thus its sensitivity is suppressed by a factor $2/5$ compared with $\langle|Q|^2\rangle =1$ case.
It remains to assign a quality factor for the third IFO. We wish to make this operate below the RMS average, % in terms of its orientation and the sky location of the sources,
but it would be pointless to simply set its $\langle|Q|^2\rangle$ equal to zero. We therefore somewhat arbitrarily set $\psi=0,\iota=0$
and introduce the following $\theta,\phi$-only averages
%
\be
 \langle F_+^2\rangle_{\theta,\phi} = \frac{7}{30},\qquad \langle F_\times^2\rangle_{\theta,\phi} = \frac{1}{6}\, 
\ee
yielding
%
\be
\langle |Q(\theta,\phi,\psi=0,\iota=0)|^2 \rangle_{\theta,\phi} = \f{2}{5} \label{eq:RMS_th_phi}.
\ee
Using this value, we assign our below-average third IFO $\langle|Q|^2\rangle^{1/2}=4/5-\sqrt{2/5}\simeq 0.1675$.
%which is essentially a $\{\theta$-$\phi\}$-RMS awayfrom the overall RMS averaged value of 2/5.
This may seem somewhat ad hoc, but it is our way of modeling the Advanced-LIGO-Virgo network of 2020s in which 
the three IFOs will accumulate varying SNRs due to their differing quality factors.
%

Therefore, to represent our multi-interferometer network, we introduce % the norm of the frequency-domain strain (\ref{eq:strain_FD}) 
%%
%
\be
\tilde{H}_i(f)\equiv \tilde{A}_i\, h_0\, f^{-7/6} \label{eq:Hi_s}
\ee
%
%%
for each interferometer, where
%%
%
\begin{align}
%\tilde{A}_i &\equiv A \langle |Q|^2\rangle^{1/2}_i=\f{1}{2}\sqrt{\f{5}{6}}\, \pi^{-2/3} \left\langle\, |Q_i(\theta,\phi,\psi,\iota)|^2\,\right\rangle^{1/2}\quad\text{and} \label{eq:A_tilde}\\
h_0 &= \f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6} \label{eq:h0}\, .
\end{align}
%
and the $\tilde{A}_i =A\langle|Q|^2\rangle^{1/2}_i $ are given in Table~\ref{table:network_params}. %for $i=1,2,3$. %Clearly, the strain and thus the SNR at each of our imaginary interferometers scales as $\langle |Q|^2\rangle^{1/2}_i$, thus
%%
The SNR for each interferometer can now be written as
%%
%
\begin{align}
\rho_i &= 2\tilde{A}_i\, h_0 \left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{n}(f)} \right]^{1/2}\label{eq:SNR_i}.
\end{align}
%
We arbitrarily make the following assignments for our network: $h=1, m=2, l=3$ representing high, medium, low strains/SNRs, respectively. 
We summarize our network in the following Table~\ref{table:network_params}.
%%

\begin{table}[h]
\centering
\begin{tabular}{c|c|c|c}
 IFO$_i$ & configuration %& $\langle |Q_i|^2\rangle$ 
 & $\tilde{A}_i$ & $\tilde{H}_i(f)$\\
 %\hline
 \hline
 1 & high ($h$) &  A & $A h_0\, f^{-7/6}$ \T\\
 2 & medium ($m$) & $\tfrac{2}{5}A$ & $ \tfrac{2}{5}A h_0 \, f^{-7/6}$\B\T \\
 3 & low ($l$)  & $\left(\tfrac{4}{5}-\sqrt{\tfrac{2}{5}}\right)A$ & $ \left(\tfrac{4}{5}-\sqrt{\tfrac{2}{5}}\right)A h_0 \, f^{-7/6}$\B \\
 \hline
\end{tabular}
\caption{%The averaged quality factors $\langle|Q|^2\rangle$ of 
Our model for the Advanced-LIGO-Virgo network ca. 2020. %of three L-shaped interferometers (IFOs). 
Column four lists the magnitude of the strain for each IFO via Eq.~(\ref{eq:Hi_s}).
``Configuration'' describes how ideally oriented/positioned an IFO is with respect to a compact binary source. 
High is $2\sigma$ above the average as explained above.
Medium is the RMS-averaged (one $\sigma$) response. Recall $A= \pi^{-2/3} \sqrt{5/24}$. 
%Note that for an interferometer where the angle between the arms is $\alpha$, $A$ rescales to $A\sin\alpha$.
}\label{table:network_params}
\end{table}
%
%
We can now write the total SNR for our network
%%
%
\begin{align}
\rho_\text{tot} &= \sqrt{\rho^2_h+\rho^2_m+\rho^2_l} \label{eq:SNR_total_3IFOs} .
%	      &= 2A h_0 \sqrt{1+\tfrac{2}{5}+\tfrac{4}{25}} \left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{noise}(f)}\right]^{1/2}\nn\\
%	      &= \sqrt{\frac{13}{10}}\,\pi^{-2/3}\,\f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6}\left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{noise}(f)}\right]^{1/2} \label{eq:SNR_total_3IFOs_v2}.
\end{align}
%
%%
%Although we made several simplifying assumptions to obtain $\rho_\text{tot}$, these should not affect our total SNR estimation too much.
%For example, averaging over all the angles underestimates the SNR of an interferometer with a near-ideal orientation with respect to the incoming GWs
%while also overestimating the SNR of a badly oriented detector. Rewriting Eq.~(\ref{eq:SNR_total}) as Eq.~(\ref{eq:SNR_total_3IFOs}) compensates
%for these.
We will start Sec.~\ref{sec:results} by testing our network with GW170817 whose signal got picked up by the LIGO detectors around $f=30$\,Hz.
So we will change the lower limit of the integral in Eq.~(\ref{eq:SNR_total_3IFOs}) from $0$ to $30$.
For our advance warning times we will further use separate initial frequencies for each inteferometer as we will explain in Sec.~\ref{sec:results}.

@@

The amplitude-adjusted strains $ \tilde{H}_i(f)$ are represented by the dashed ($i=h$), solid ($i=m$), and dotted ($i=l$) lines in the figure where we actually plot
$2\sqrt{f} \tilde{H}_i(f)$ as is standard. 

@@

\bar\rho_h = 29.2,\quad \bar\rho_m = 14.6,\quad \bar\rho_l = 6.11 \label{eq:GW170817_SNRs}.

@@

%To accommodate more realistic orientations, we show two shaded regions in Fig.~\ref{fig:figGW170817}. The inspirals in the lighter region yield SNR$\ge 22.34$ as
%they sweep from 10\,Hz to $f_\text{ISCO} \simeq1605\,$Hz whereas inspirals in the darker region accumulate $ \bar\rho_l \le \text{SNR} \le 22.34$.
The shaded region in Fig.~\ref{fig:figGW170817} covers strains between $\tilde{H}_l(f)$ and $\tilde{H}_h(f)$. %which we expect to represent mo
Given a network of three equally sensitive, randomly oriented IFOs we expect that a single BNS inspiral would produce three distinct detector
response strains in this shaded region.

@@

\caption{GW170817 as it sweeps from 10\,Hz to $f_\text{ISCO} \simeq1605\,$Hz across our network with ASD (sensitivity) of 
LIGO Livingstone (L1) during the event (the darker red curve).
We also show the Hanford detector's sensitivity in light blue for comparison. The dashed, solid and dotted lines represent the 
response of each interferometer's frequency-domain response ($2\sqrt{f} \tilde{H}_i(f)$ with $i=h,m,l$) to the inspiral of
a binary neutron star system with parameters given in Eq.~(\ref{eq:GW170817_params}).
The variation in strain amplitudes is the result of a given interferometer's orientation and the sky position of the source 
as explained in Sec.~\ref{sec:LIGO_topo} and summarized in Table~\ref{table:network_params}.
In other words, the dashed line is how ``loud'' GW170817 would appear in the data stream of a near optimally oriented IFO with L1's sensitivity

@@

(quality factor $Q=4/5$). Similarly, the solid line represents the strain in %a randomly oriented IFO with randomized source sky location, hence
the RMS-averaged IFO with L1's sensitivity ($Q=2/5$). %, but with RMS-average over all angles 

@@

resulting in a reduction of amplitude to $ 16.7\%$ of the $Q=1$ case. 

@@

network, we would expect all three strains to exist somewhere in the shaded region most of the time and two out of three every time.

@@

We can now compute the total SNR of GW170817 as inferred by our network
%%
%
\be
\bar\rho_\text{tot}= \sqrt{\bar\rho^2_h+\bar\rho^2_m+\bar\rho^2_l}= 32.6 \, \label{eq:SNR_total_GW170817} .
\ee
%
%%
This also compares well with the actual value of 32.4. Therefore, we conclude that our imaginary network performs realistically enough to make predictions on the performance of IFOs in the 2020s and 2030s.
%
%
%
%
%
\section{Results: Advance warning times for inspiralling neutron stars}\label{sec:results}
 %\subsection{BNS inspirals at Advanced LIGO-Virgo and the Einstein Telescope}\label{Sec:BNS_2020_2030}
 %Recall that we are interested in the advance warning times that the (near)-future ground-based interferometers can provide for the optical astronomers before the prompt GRB at the end of the merger. 
 Our aim is to foretell GRBs using our network made up of three IFOs with A-LIGO design sensitivity in the 2020s and the Einstein Telescope sensitivity in the 2030s.
 Our predictive capabilities depend on how much advance warning the future networks will give us prior to the prompt merger/GRB.
 Our definition of advance warning is the inspiral time to the merger from a certain threshold instant $\bar{t}$ at which point the network ``agrees'' that there is an inspiral 
 event statistically significant enough to issue warnings to electromagnetic telescopes.
 We define this threshold instant to be given at a frequency $\bar{f}$ when the total SNR equals 15. 
 More specifically, in the case of the Advanced-LIGO-Virgo network (ALV), our statement is that there exists a frequency $\bar{f} < f_\text{ISCO}$ such that
 %
 \be
 \bar\rho_\text{tot}= \sqrt{\left[\rho_h(f_0^h,\bar{f})\right]^2+\left[\rho_m(f_0^m,\bar{f})\right]^2+\left[\rho_l(f_0^l,\bar{f})\right]^2}=15, \label{eq:SNR_bar}%2\tilde{A}_m\, h_0\left[ \int_{f_0}^{{\bar{f}}} df'\, \f{f'^{-7/3}}{S_\text{noise}(f')}\right]^{1/2} =15 .
 \ee
 %
 where $f_0^i$ are given by the solution set to
 %there exists some frequency $f_0 < \bar{f}$ such that
 \be
 \sqrt{S_\text{n}(f_0^i)} = 2\sqrt{f_0^i}\, \tilde{H}_i(f_0^i) \label{eq:f0},\qquad i=h,m,l
 \ee
 %
 with $f_0^i < \bar{f}$. In other words, $f_0^i$ are the initial GW frequencies at which a given inspiral enters the $i$th IFO's sensitivity band. 
 Once we know $\bar{f}$, we can compute the remaining time to the merger, $\tau_\text{insp}(\bar{f})$, using Eq.~(\ref{eq:tau_insp})
 and its 3.5PN-enhanced version from Ref.~\cite{Blanchet_LRR}. This is what we call our advance warning time $T_\text{AW}$.
 For simplicity, we set $m_1=m_2=1.4 M_\odot \implies M_c\simeq 1.219 M_\odot$, which we justified in
 Point 3 of Sec.~\ref{sec:idealizations}. 
 With the masses fixed, the only variable determining $h_0$ is the luminosity distance $D$ as can be seen from Eq.~(\ref{eq:h0}).
 We summarize our procedure in the flow diagram below.
 %
 %
  \begin{figure}[h!]
 \framebox{\Longstack[l]{Pick $D$}}$\longrightarrow$\framebox{\Longstack[l]{Construct\\ $\tilde{H}_i(f)$}} 
 $\longrightarrow$  \framebox{\Longstack[l]{Compute $f_0^i$\\ from Eq.(\ref{eq:f0})}} $\longrightarrow$  \framebox{\Longstack[l]{Compute $\bar{f}$\\ from Eq.(\ref{eq:SNR_bar})}}
 $\longrightarrow$  \framebox{\Longstack[l]{Compute $T_{AW}=\tau_\text{insp}(\bar{f})$\\ from Eq.~(\ref{eq:tau_insp})}}
  \caption{Flow diagram summarizing our procedure for computing advance warning times $T_\text{AW}$ with the Advanced-LIGO-Virgo network. 
  We will compute $T_\text{AW}$ for the case of $M=2.8M_\odot$ BNS system inspiralling at $D=50,100,200\,$Mpc. 
  %for A-LIGO and $D=200,400,1000\,$Mpc for the ET.
  }\label{fig:flow}
\end{figure}

@@

\section{Binary black hole neutron star systems}
In this section, we investigate the possibility of witnessing the tidal distruption of a neutron star as it falls into a black hole. 
The treatment of the inspiral is the same as before, however the fate of the neutron star depends strongly on the system's intrinsic parameters
such as the black hole mass and spin, and the neutron star's equation of state (EOS) and compactness $C_\text{NS}\equiv G m_\text{NS}/(R_\text{NS}c^2)$. 
The evolution of black hole-neutron star (BH-NS) binaries is a very active branch of numerical relativity
(see Ref. \cite{Shibata:2011jka} for a comprehensive review) and is beyond the scope of this article, however the numerical simulaitions
have verified certain qualitative principles that will guide us:{\bf fix this sentence}
%
\begin{enumerate}[(i)]
 \item Small separations. The Newtonian tidal force due to the black hole (BH) is proportional to $M_\text{BH}/r^3$. Therefore, $r$ has to be
 minimized to ensure maximum tide.
 \item ISCO as a lower bound. However, for $r< r_\text{ISCO}$ the NS simply plunges into the BH; therefore the minium separation is given by the ISCO radius.
 \item Small black hole masses. In the geodecis case, $r_\text{ISCO} \propto M_\text{BH}$, so the Newtonian tidal force roughly scales as $M_\text{BH}^{-2}$. 
 \item High black hole spins. Moreover, $r_\text{ISCO}$ is much smaller in the case of a prograde NS orbit around a spinning Kerr black hole.
 For instance, for the maximum spin value of $a=1$, $r_\text{ISCO}$ for prograde orbits is 1/6th of the Schwarzschild value whereas for
 $a=1$ retrograde orbits, $r_\text{ISCO}$ is 3/2 of the Schwarzschild value.
\end{enumerate}

We will set $a=0.9$ cite, $M_\text{BH} =5 M_\odot$ cite(Morgan's email)

The less massive BHs are also desirable as 
 $\tau_\text{insp} \propto M_c^{-5/3} = (1+Q)^{1/3}/Q$ for $Q\equiv m_2/m_1 > 1$.
